---
title: "Milestone 8"
author: "Jenna Moustafa"
date: "4/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 15)
library(tidyverse)
library(readxl)
library(janitor)
library(tm)
library(writexl)
library(broom)
library(gt)

# Read in fourth grade reading scores by state and year, keeping only
# observation rows

fourth_reading <- read_xlsx(path = "raw-data/fourth_reading_naep.xlsx",
                            skip = 7, n_max = 531) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth") %>%
  mutate(subject = "reading") %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "average_scale_score")

# Read in fourth grade math scores by state and year, keeping only
# observation rows

fourth_math <- read_xlsx(path = "raw-data/fourth_math_naep.xlsx",
                         skip = 7, n_max = 531) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth") %>%
  mutate(subject = "math") %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "average_scale_score")

# Read in eighth grade reading scores by state and year, keeping only
# observation rows

eighth_reading <- read_xlsx(path = "raw-data/eighth_reading_naep.xlsx",
                            skip = 7, n_max = 531) %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth") %>%
  mutate(subject = "reading") %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "average_scale_score")

# Read in eighth grade math scores by state and year, keeping only
# observation rows

eighth_math <- read_xlsx(path = "raw-data/eighth_math_naep.xlsx",
                         skip = 7, n_max = 531) %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth") %>%
  mutate(subject = "math") %>%
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "average_scale_score")

# Read in high school graduation rates by state and year, keeping only
# observation rows for years graduating years 2011, 2012, and 2013
# replacing dash with NA, and renaming columns so they are not just numeric

grad_rate_2011_2013 <- read_excel("raw-data/grad_rate_2011_2013.xlsx",
                                  col_names = c("state", "grad_2011",
                                                "grad_2012", "notes",
                                                "grad_2013"),
                                  na = "—", skip = 5, n_max = 52) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Remove column with useless info
  
  select("state", "grad_2011", "grad_2012", "grad_2013") %>% 
  
  # Fix state cells that have numbers included from footnotes
  
  mutate(state = removeNumbers(state))
```

```{r, echo = FALSE}
# Join fourth grade reading and math scores together so that each table is on 
# top of each other
fourth <- fourth_reading %>% 
  full_join(fourth_math, by = c("year", "jurisdiction", "grade", "subject",
                                "average_scale_score"))

# Join eighth grade reading and math scores together so that each table is on 
# top of each other

eighth <- eighth_reading %>% 
  full_join(eighth_math, by = c("year", "jurisdiction", "grade", "subject",
                                "average_scale_score"))

# Join eighth grade and fourth grade scores together so that all tables are on 
# top of each other

naep <- fourth %>% 
  full_join(eighth,  by = c("year", "jurisdiction", "grade", "subject",
                                "average_scale_score")) %>% 
  mutate(score = as.numeric(average_scale_score)) %>% 
  mutate(test_type = case_when(grade == "fourth" &
                                 subject == "reading" ~ "fourth reading",
                               grade == "fourth" &
                                 subject == "math" ~ "fourth math",
                               grade == "eighth" &
                                 subject == "reading" ~ "eighth reading",
                               grade == "eighth" &
                                 subject == "math" ~ "eighth math")) %>% 
  mutate(cc_implementation = case_when(jurisdiction == "Alabama" ~ 2013,
                             jurisdiction == "Arizona" ~ 2013,
                             jurisdiction == "Arkansas" ~ 2013,
                             jurisdiction == "California" ~ 2014,
                             jurisdiction == "Colorado" ~ 2013,
                             jurisdiction == "Connecticut" ~ 2013,
                             jurisdiction == "Delaware" ~ 2012,
                             jurisdiction == "Georgia" ~ 2014,
                             jurisdiction == "Hawaii" ~ 2013,
                             jurisdiction == "Idaho" ~ 2013,
                             jurisdiction == "Illinois" ~ 2013,
                             jurisdiction == "Iowa" ~ 2012,
                             jurisdiction == "Kansas" ~ 2013,
                             jurisdiction == "Kentucky" ~ 2011,
                             jurisdiction == "Louisiana" ~ 2013,
                             jurisdiction == "Maine" ~ 2012,
                             jurisdiction == "Maryland" ~ 2013,
                             jurisdiction == "Massachusetts" ~ 2013,
                             jurisdiction == "Michigan" ~ 2012,
                             jurisdiction == "Mississippi" ~ 2013,
                             jurisdiction == "Missouri" ~ 2014,
                             jurisdiction == "Montana" ~ 2013,
                             jurisdiction == "New Hampshire" ~ 2014,
                             jurisdiction == "New Jersey" ~ 2013,
                             jurisdiction == "New Mexico" ~ 2013,
                             jurisdiction == "New York" ~ 2013,
                             jurisdiction == "North Carolina" ~ 2012,
                             jurisdiction == "North Dakota" ~ 2013,
                             jurisdiction == "Ohio" ~ 2013,
                             jurisdiction == "Oregon" ~ 2014,
                             jurisdiction == "Pennsylvania" ~ 2013,
                             jurisdiction == "Rhode Island" ~ 2013,
                             jurisdiction == "South Dakota" ~ 2014,
                             jurisdiction == "Tennessee" ~ 2013,
                             jurisdiction == "Utah" ~ 2013,
                             jurisdiction == "Vermont" ~ 2013,
                             jurisdiction == "Washington" ~ 2014,
                             jurisdiction == "West Virginia" ~ 2014,
                             jurisdiction == "Wisconsin" ~ 2013,
                             jurisdiction == "Wyoming" ~ 2014),
         treat_year = ifelse(year < cc_implementation, 0, 1),
         treat_year = replace_na(treat_year, 0),
         test_type = case_when(test_type == "fourth reading" ~ "Fourth Grade Reading",
                               test_type == "eighth reading" ~ "Eighth Grade Reading",
                               test_type == "fourth math" ~ "Fourth Grade Math",
                               test_type == "eighth math" ~ "Eighth Grade Math"))
```

```{r}
### Repeat above steps for subgroups in each grade and subject
## Fourth grade reading scores
## ethnicity

fourth_reading_race <- read_xlsx(path = "raw-data/fourth_reading_race.xlsx",
                            skip = 8, n_max = 3189-8) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
    
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>% 
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth",
         subject = "reading",
         score = as.numeric(average_scale_score)) %>% 
  
  # Rename subgroup column name

  rename(race = race_ethnicity_used_to_report_trends_school_reported) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "race", "score")

## Ethnicity

fourth_reading_ethn <- read_xlsx(path = "raw-data/fourth_reading_ethn.xlsx",
                         skip = 8, n_max = 3719-8) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth",
         subject = "reading",
         score = as.numeric(average_scale_score)) %>%  
  
  # Rename subgroup column name

  rename(ethnicity = race_ethnicity_using_2011_guidelines_school_reported) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "ethnicity", "score")

## Gender

fourth_reading_gender <- read_xlsx(path = "raw-data/fourth_reading_gender.xlsx",
                            skip = 8, n_max = 1069-8) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth",
         subject = "reading",
         score = as.numeric(average_scale_score)) %>%  
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "gender", "score")

# Read in fourth grade math scores by state and year, keeping only
# observation rows

fourth_reading_lunch <- read_xlsx(path = "raw-data/fourth_reading_lunch.xlsx",
                         skip = 8, n_max = 1599-8) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth",
         subject = "reading",
         score = as.numeric(average_scale_score)) %>%  
  
  # Rename subgroup column name
  
  rename(school_lunch = national_school_lunch_program_eligibility_3_categories) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "school_lunch", "score")

# Join fourth grade reading subgroup scores together so that each table is on 
# top of each other

fourth_reading_sub <- merge(fourth_reading_race, fourth_reading_gender, by = c("year", "jurisdiction", "grade", "subject",
                                "score"), all = TRUE) %>% 
  merge(., fourth_reading_lunch, by = c("year", "jurisdiction", "grade", "subject",
                                "score"), all = TRUE) %>% 
  filter(!is.na(score)) %>% 
  select(-score, everything())
```


```{r}
### Repeat above steps for subgroups in each grade and subject
## Fourth grade math scores
## ethnicity

fourth_math_race <- read_xlsx(path = "raw-data/fourth_math_race.xlsx",
                            skip = 8, n_max = 3189-8) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth",
         subject = "math",
         score = as.numeric(average_scale_score)) %>% 
  
  # Rename subgroup column name

  rename(race = race_ethnicity_used_to_report_trends_school_reported) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "race", "score")

## Ethnicity

fourth_math_ethn <- read_xlsx(path = "raw-data/fourth_math_ethn.xlsx",
                         skip = 8, n_max = 3719 - 8) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth",
         subject = "math",
         score = as.numeric(average_scale_score)) %>%  
  
  # Rename subgroup column name

  rename(ethnicity = race_ethnicity_using_2011_guidelines_school_reported) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "ethnicity", "score")

## Gender

fourth_math_gender <- read_xlsx(path = "raw-data/fourth_math_gender.xlsx",
                            skip = 8, n_max = 1069-8) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth",
         subject = "math",
         score = as.numeric(average_scale_score)) %>%  
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "gender", "score")

# Read in fourth grade math scores by state and year, keeping only
# observation rows

fourth_math_lunch <- read_xlsx(path = "raw-data/fourth_math_lunch.xlsx",
                         skip = 8, n_max = 1599-8) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "fourth",
         subject = "math",
         score = as.numeric(average_scale_score)) %>%  
  
  # Rename subgroup column name
  
  rename(school_lunch = national_school_lunch_program_eligibility_3_categories) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "school_lunch", "score")

# Join fourth grade reading subgroup scores together so that each table is on 
# top of each other

fourth_math_sub <- merge(fourth_math_race, fourth_math_gender, by = c("year", "jurisdiction", "grade", "subject",
                                "score"), all = TRUE) %>% 
  merge(., fourth_math_lunch, by = c("year", "jurisdiction", "grade", "subject",
                                "score"), all = TRUE) %>% 
  filter(!is.na(score)) %>% 
  select(-score, everything())
```


```{r}
### Repeat above steps for subgroups in each grade and subject
## Eighth grade reading scores
## ethnicity

eighth_reading_race <- read_xlsx(path = "raw-data/eighth_reading_race.xlsx",
                            skip = 8, n_max = 3507-8) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth",
         subject = "reading",
         score = as.numeric(average_scale_score)) %>% 
  
  # Rename subgroup column name

  rename(race = race_ethnicity_used_to_report_trends_school_reported) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "race", "score")

## Ethnicity

eighth_reading_ethn <- read_xlsx(path = "raw-data/eighth_reading_ethn.xlsx",
                         skip = 8, n_max = 4090-8) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth",
         subject = "reading",
         score = as.numeric(average_scale_score)) %>%  
  
  # Rename subgroup column name

  rename(ethnicity = race_ethnicity_using_2011_guidelines_school_reported) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "ethnicity", "score")

## Gender

eighth_reading_gender <- read_xlsx(path = "raw-data/eighth_reading_gender.xlsx",
                            skip = 8, n_max = 1175-8) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth",
         subject = "reading",
         score = as.numeric(average_scale_score)) %>%  
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "gender", "score")

## Lunch

eighth_reading_lunch <- read_xlsx(path = "raw-data/eighth_reading_lunch.xlsx",
                         skip = 8, n_max = 1758-8) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth",
         subject = "reading",
         score = as.numeric(average_scale_score)) %>%  
  
  # Rename subgroup column name
  
  rename(school_lunch = national_school_lunch_program_eligibility_3_categories) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "school_lunch", "score")

# Join fourth grade reading subgroup scores together so that each table is on 
# top of each other

eighth_reading_sub <- merge(eighth_reading_race, eighth_reading_gender, by = c("year", "jurisdiction", "grade", "subject",
                                "score"), all = TRUE) %>% 
  merge(., eighth_reading_lunch, by = c("year", "jurisdiction", "grade", "subject",
                                "score"), all = TRUE) %>% 
  filter(!is.na(score)) %>% 
  select(-score, everything())
```


```{r}
### Repeat above steps for subgroups in each grade and subject
## Eighth grade math scores
## ethnicity

eighth_math_race <- read_xlsx(path = "raw-data/eighth_math_race.xlsx",
                            skip = 8, n_max = 3189-8) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth",
         subject = "math",
         score = as.numeric(average_scale_score)) %>% 
  
  # Rename subgroup column name

  rename(race = race_ethnicity_used_to_report_trends_school_reported) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "race", "score")

## Ethnicity

eighth_math_ethn <- read_xlsx(path = "raw-data/eighth_math_ethn.xlsx",
                         skip = 8, n_max = 3719-8) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth",
         subject = "math",
         score = as.numeric(average_scale_score)) %>%  
  
  # Rename subgroup column name

  rename(ethnicity = race_ethnicity_using_2011_guidelines_school_reported) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "ethnicity", "score")

## Gender

eighth_math_gender <- read_xlsx(path = "raw-data/eighth_math_gender.xlsx",
                            skip = 8, n_max = 1069-8) %>%
  
  # Make column names easy to use, so that there are no spaces
  
  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth",
         subject = "math",
         score = as.numeric(average_scale_score)) %>%  
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "gender", "score")

# Read in fourth grade math scores by state and year, keeping only
# observation rows

eighth_math_lunch <- read_xlsx(path = "raw-data/eighth_math_lunch.xlsx",
                         skip = 8, n_max = 1599-8) %>%
  
  # Make column names easy to use, so that there are no spaces

  clean_names() %>%
  
  # Remove NA observations
  
  filter(!is.na(year) & jurisdiction != "DoDEA" & jurisdiction != "District of Columbia") %>%
  
  # Replace symbols with NA
  
  na_if(.,"‡") %>%
  na_if(., "—") %>%
  
  # Create columns for grade and subjects
  
  mutate(grade = "eighth",
         subject = "math",
         score = as.numeric(average_scale_score)) %>%  
  
  # Rename subgroup column name
  
  rename(school_lunch = national_school_lunch_program_eligibility_3_categories) %>% 
  
  # Remove column with useless info and reorganize column so score is last
  
  select("year", "jurisdiction", "grade", "subject", "school_lunch", "score")

# Join fourth grade reading subgroup scores together so that each table is on 
# top of each other

eighth_math_sub <- merge(eighth_math_gender, eighth_math_race,
                         by = c("year", "jurisdiction", "grade",
                                "subject", "score"), all = TRUE) %>%
  merge(., eighth_math_lunch, by = c("year", "jurisdiction", "grade",
                                "subject", "score"), all = TRUE) %>%
  filter(!is.na(score)) %>% 
  select(-score, everything())

```

```{r, echo = FALSE}
# Join fourth grade reading and math scores together so that each table is on 
# top of each other
fourth_sub <- fourth_reading_sub %>% 
  merge(., fourth_math_sub, by = c("year", "jurisdiction", "grade", "subject",
                                "score", "race", "gender",
                                "school_lunch"), all = TRUE)

# Join eighth grade reading and math scores together so that each table is on 
# top of each other

eighth_sub <- eighth_reading_sub %>% 
  merge(., eighth_math_sub, by = c("year", "jurisdiction", "grade", "subject",
                                "score", "race", "gender",
                                "school_lunch"), all = TRUE)

# Join eighth grade and fourth grade scores together so that all tables are on 
# top of each other

naep_sub <- fourth_sub %>% 
  merge(., eighth_sub,  by = c("year", "jurisdiction", "grade", "subject",
                                "score", "race", "gender",
                                "school_lunch"), all = TRUE) %>%
  mutate(test_type = case_when(grade == "fourth" &
                                 subject == "reading" ~ "fourth reading",
                               grade == "fourth" &
                                 subject == "math" ~ "fourth math",
                               grade == "eighth" &
                                 subject == "reading" ~ "eighth reading",
                               grade == "eighth" &
                                 subject == "math" ~ "eighth math")) %>% 
  select(-score, everything()) %>% 
  mutate(cc_implementation = case_when(jurisdiction == "Alabama" ~ 2013,
                             jurisdiction == "Arizona" ~ 2013,
                             jurisdiction == "Arkansas" ~ 2013,
                             jurisdiction == "California" ~ 2014,
                             jurisdiction == "Colorado" ~ 2013,
                             jurisdiction == "Connecticut" ~ 2013,
                             jurisdiction == "Delaware" ~ 2012,
                             jurisdiction == "Georgia" ~ 2014,
                             jurisdiction == "Hawaii" ~ 2013,
                             jurisdiction == "Idaho" ~ 2013,
                             jurisdiction == "Illinois" ~ 2013,
                             jurisdiction == "Iowa" ~ 2012,
                             jurisdiction == "Kansas" ~ 2013,
                             jurisdiction == "Kentucky" ~ 2011,
                             jurisdiction == "Louisiana" ~ 2013,
                             jurisdiction == "Maine" ~ 2012,
                             jurisdiction == "Maryland" ~ 2013,
                             jurisdiction == "Massachusetts" ~ 2013,
                             jurisdiction == "Michigan" ~ 2012,
                             jurisdiction == "Mississippi" ~ 2013,
                             jurisdiction == "Missouri" ~ 2014,
                             jurisdiction == "Montana" ~ 2013,
                             jurisdiction == "New Hampshire" ~ 2014,
                             jurisdiction == "New Jersey" ~ 2013,
                             jurisdiction == "New Mexico" ~ 2013,
                             jurisdiction == "New York" ~ 2013,
                             jurisdiction == "North Carolina" ~ 2012,
                             jurisdiction == "North Dakota" ~ 2013,
                             jurisdiction == "Ohio" ~ 2013,
                             jurisdiction == "Oregon" ~ 2014,
                             jurisdiction == "Pennsylvania" ~ 2013,
                             jurisdiction == "Rhode Island" ~ 2013,
                             jurisdiction == "South Dakota" ~ 2014,
                             jurisdiction == "Tennessee" ~ 2013,
                             jurisdiction == "Utah" ~ 2013,
                             jurisdiction == "Vermont" ~ 2013,
                             jurisdiction == "Washington" ~ 2014,
                             jurisdiction == "West Virginia" ~ 2014,
                             jurisdiction == "Wisconsin" ~ 2013,
                             jurisdiction == "Wyoming" ~ 2014),
         treat_year = ifelse(year < cc_implementation, 0, 1),
         treat_year = replace_na(treat_year, 0),
         non_white = ifelse(race == "White", 0, 1),
         black = ifelse(race == "Black", 1, 0),
         asian = ifelse(race == "Asian/Pacific Islander", 1, 0),
         hispanic = ifelse(race == "Hispanic", 1, 0),
         female = ifelse(gender == "female", 1, 0),
         low_income = ifelse(school_lunch == "Eligible", 1, 0),
         test_type = case_when(test_type == "fourth reading" ~ "Fourth Grade Reading",
                               test_type == "eighth reading" ~ "Eighth Grade Reading",
                               test_type == "fourth math" ~ "Fourth Grade Math",
                               test_type == "eighth math" ~ "Eighth Grade Math"))
```

```{r}
## Save datasets into a single file, replacing old files saved 
write_xlsx(naep_sub, path = "shiny/naep_sub.xlsx")
write_xlsx(naep, path = "shiny/naep.xlsx")
```



## Background
The goal of this project is to determine how implementation of the Common Core
Standards (CSS) in the early 2010s affected achievement relative across states
relative to states that did not implement the CSS. The goal of the curriculum
standardization is to narrow achievement gaps; however, there is debate over
whether the Common Core actually achieved this goal as well as concern over
states simply adjusting their state standardized tests to allow for teachers' 
teaching to the test. To avoid this bias, I am using National Assessment of 
Educational Progress (NAEP) data, which includes average student test scores by
states on national standardized exams.

## Data
The data from this project is from http://www.corestandards.org and 
nationsreportcard.gov, which respectively have data on dates regarding CSS
adoption and implementation across states and data on test scores by state and
year for fourth and eighth grade NAEP scores as well as by subcategory (e.g.
race, gender, school lunch program eligibility). School lunch eligibility is 
used as a proxy for low-income, since income levels are not reported/provided.

## Methodology
This project examines NAEP score trends across states, by race, gender, and 
income in order to determine how implementation of the CSS impacted
standardized exam scores for students overall and how the program impacted
achievement gaps. The project conducts an event-study design regressing
each average test score by state, subject, and grade against the nine
states that chose not to implement the CSS as controls. 
